---
title: "Getting started with shinycapture"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with shinycapture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Why browser-side capture?

Plotly charts in Shiny are rendered by JavaScript in the user's browser —
the R server never holds the chart as pixels. Standard server-side tools like
`ggplot2::ggsave()` cannot capture a Plotly figure. While headless-browser
approaches (`webshot2`, `chromote`) are possible, they re-render the chart from
scratch in a separate process and require a server-side Chrome installation.

shinycapture triggers the capture directly in the running browser and sends the
result back to R — no server-side Chrome required.

## Basic usage

```{r}
library(shiny)
library(plotly)
library(shinycapture)

ui <- fluidPage(
  plotlyOutput("my_plot"),
  actionButton("capture_btn", "Export PNG"),
  uiOutput("preview")
)

server <- function(input, output, session) {
  output$my_plot <- renderPlotly({
    plot_ly(x = ~rnorm(100), type = "histogram")
  })

  # Step 1: trigger capture in the browser
  observeEvent(input$capture_btn, {
    capture_plotly("my_plot",
      strategy = plotly_strategy(width = 1200, height = 800)
    )
  })

  # Step 2: react when the result arrives as a base64 data-URI string
  observeEvent(input[[".shinycapture.my_plot"]], {
    # Display in the UI — use the base64 string directly as img src
    output$preview <- renderUI(
      tags$img(src = input[[".shinycapture.my_plot"]])
    )
  })
}

shinyApp(ui, server)
```

The capture result arrives in `input[[".shinycapture.<id>"]]` as a base64
data-URI string. Use it directly as an `<img>` `src` for display, or decode
it to raw bytes with `base64_decode()` for saving or further processing.

## Saving to disk

```{r}
observeEvent(input[[".shinycapture.my_plot"]], {
  bytes <- base64_decode(input[[".shinycapture.my_plot"]])
  writeBin(bytes, "export.png")
})
```

## Offering a download button

```{r}
captured_bytes <- reactiveVal(NULL)

observeEvent(input$capture_btn, {
  capture_plotly("my_plot", input_id = ".shinycapture.dl")
})

observeEvent(input[[".shinycapture.dl"]], {
  captured_bytes(base64_decode(input[[".shinycapture.dl"]]))
  output$download_ui <- renderUI(downloadButton("dl", "Download PNG"))
})

output$dl <- downloadHandler(
  filename = function() "chart.png",
  content  = function(file) writeBin(captured_bytes(), file)
)
```

## Capture strategies

### Plotly — exact resolution with strip patches

`plotly_strategy()` uses `Plotly.toImage()` for pixel-perfect output.
Strip flags remove chart decorations before capture without affecting
the live chart — they operate on a hidden off-screen clone.

```{r}
capture_plotly("my_plot",
  strategy = plotly_strategy(
    strip_title  = TRUE,   # remove title
    strip_legend = TRUE,   # remove legend
    width  = 2400,
    height = 1600,
    format = "png"         # or "jpeg", "webp", "svg"
  )
)
```

### html2canvas — any HTML element

`html2canvas_strategy()` can capture any DOM element, not just Plotly charts.
The `scale` parameter controls resolution: `2` (default) gives double density,
suitable for retina displays; `3` for print quality.

```{r}
capture_element("my_table_div",
  strategy = html2canvas_strategy(scale = 3)
)
```

## Custom input IDs

By default the result lands in `input[[".shinycapture.<id>"]]`. Pass
`input_id` to use a different name — useful when capturing the same element
multiple times with different strategies.

```{r}
capture_plotly("my_plot",
  strategy = plotly_strategy(format = "svg"),
  input_id = "my_svg_capture"
)

observeEvent(input$my_svg_capture, {
  bytes <- base64_decode(input$my_svg_capture)
  writeBin(bytes, "chart.svg")
})
```

## How it works

1. `capture_plotly()` / `capture_element()` sends a custom Shiny message to the browser.
2. The JavaScript handler runs the strategy: optional preprocess step (strip patches),
   then `Plotly.toImage()` or `html2canvas`.
3. The resulting base64 string is sent back via `Shiny.setInputValue()`.
4. Your `observeEvent` fires and the string is available in `input[[id]]`.
